A Hybrid Ensemble Learning Approach to Star-galaxy Classification

### 2015-04-15 paper/main.tex
fix purity vs mag figure; add purity vs z.
update up to figure 8.

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[letterpaper,margin=0.65in]{geometry}
\usepackage[ansinew]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
\usepackage{amssymb}
\usepackage[normalem]{ulem}
\usepackage{natbib}
\usepackage{url}

@@

\label{firstpage}
\begin{abstract}
\textcolor{red}{Abstract goes here.}
\end{abstract}

@@

In Figure~\ref{fig:purity_mag}, 
we compare the star and galaxy purity values
for morphological separation, TPC, and BMC techniques
as a function of $i$-band magnitude.
Although morphological separation shows a slightly better performance
at bright magnitudes $i < 21$,
BMC outperforms both morphological separation and TPC
at faint magnitudes $i > 21$.
As shown in the histogram, the histogram peaks at $i \sim 22$,
and BMC therefore outperforms the morphological separation technique
for the majority of objects.
It is also clear that BMC outperforms TPC over all magnitudes.
BMC can accomplish this by combining information from
all base classifiers,
\eg giving more weight to the morphological separation method
at bright magnitudes.

\begin{figure}
%%  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
    \caption{Purity of morphological separation,
             TPC, and BMC techniques in 
             Table~\ref{table:metrics_vvds}
             as a function of the $i$-band magnitude.
             The top panel shows the histogram of objects in the test data set.
             The second panel shows the fraction of stars
             as a function of magnitude. The error bars are Poisson.
             The bottom two panels compares
             the galaxy and star purity for morphological separation, TPC,
             and BMC techniques as a function of magnitude.
             The error bars for completeness and purity are not shown for
             clarity, but they are approximately 10 percent.
             }.
    \label{fig:purity_mag}
\end{figure}

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
    \caption{The classification performance metrics for TPC, HB, morphology,
      and BMC as applied to the CFHTLenS data in the VVDS field
      with various magnitude cuts.
      The top panel shows the number of sources in the training set
      at corresponding magnitude cuts.
      We show only one of the four combination methods, BMC,
      which has the best overall performance.}
    \label{fig:perform_mag_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
      $z_{\text{spec}}$ cuts.}
    \label{fig:perform_z_cut}
  \end{minipage}
\end{figure*}

\subsection{Insufficient Training Data}
  \label{section:poor_training}

We also investigate the impact of training set quality
by analyzing in detail a more realistic case
where the training data is available
only for a small number of objects at bright magnitudes.
To emulate this scenario,
we only use objects that have spectroscopic labels
from the VVDS 0226-04 field (which is located within the CFHTLS W1 field)
and impose a magnitude cut of $i < 22.0$ in the training data,
leaving us a training set with 1,365 objects.

\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           as applied to the CFHTLenS data in the VVDS field
           with a magnitude cut of $i < 22.0$.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
  \begin{tabular}{l c c c c c c}
    Classifier & AUC & MSE &
    $p_{g}\left(c_g=0.9964\right)$ & $p_{s}\left(c_s=0.7145\right)$ &
    $p_{g}\left(c_g=0.9600\right)$ & $p_{s}\left(c_s=0.2500\right)$\\
    \hline
    TPC  & 0.9601 & 0.0728 & 0.9438 & 0.9329 & 0.9819 & 0.9950\\
    SOMc & 0.9008 & 0.1679 & 0.8843 & 0.5044 & 0.9244 & 0.7726\\
    HB   & 0.8960 & 0.0894 & 0.9074 & 0.7516 & 0.9615 & 0.7793\\
    Morphology   & -      & \textbf{0.0397} & 0.9597 & 0.9666 & - & -\\
    WA   & 0.9639 & 0.0581 & 0.9613 & 0.9787 & 0.9756 & 1.0000\\
    BoM  & 0.9601 & 0.0728 & 0.9438 & 0.9329 & 0.9819 & 0.9950\\
    Stacking  & 0.9644 & 0.0613 & 0.9633 & 0.9835 & 0.9840 & 1.0000\\
    BMC  & \textbf{0.9738} & 0.0437 & \textbf{0.9676} & \textbf{0.9942} &
    \textbf{0.9878} & 1.0000\\
  \end{tabular}
  \label{table:metrics_vvds}
\end{table*}

We apply the same four star-galaxy classification techniques
and four different combination methods,
and test them on the same test data.
We present in Table~\ref{table:metrics_vvds}
the same six metrics for each method,
and highlight the best method for each metric.
Overall, the results obtained for the reduced data set are remarkable.
With poor training data, our data-driven methods,
\texttt{TPC} and \texttt{SOMc},
suffer a significant decrease in performance,
although they still in general perform better than 
the template fitting algorithm
\texttt{HB}.
The performance of our morphological separation method
does not depend on the training data,
and shows the best result in the MSE metric.
It also outperforms other base classifiers
in galaxy purity at galaxy completeness of 0.9964
and in star purity at star completeness of 0.7145.

@@

%% ROC curve for vvds i < 20
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/roc_cut.eps}
  \caption{A magnified view of the ROC curve of
           four star-galaxy individual classification approaches
           (TPC, SOMc, HB, and morphology)
           and the best-performing combination technique (BMC).
           trained with the reduced data set.
           We show only one of the four combination methods, BMC,
           which has the best overall performance.}
  \label{fig:roc_cut}
\end{figure}

In Figure~\ref{fig:roc_cut}, we plot the ROC curve
of all four individual classification algorithm
and our BMC technique for the poor training data case.
We note that a binary classifier,
such as our cut-based morphological separation,
whose output is either zero or one,
is a single point in the ROC space.
We vary the half-light radius to create a smooth ROC curve
for morphological separation.
It is clear that our BMC technique dominates
each individual classification techniques in the ROC space.

%% weights map for vvds i < 20
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/weights_cut.eps}
  \caption{A two-dimensional 10$\times$10 SOM representation
           showing the relative weights for the BMC combination technique
           applied to the four individual methods
           trained with the CFHTLenS data in the VVDS field
           with $i < 22$.}
  \label{fig:weights_cut}
\end{figure}

@@

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_mag_cut.eps}
    \caption{Purity of morphological separation,
             TPC, and BMC techniques in 
             Table~\ref{table:metrics_vvds}
             as a function of the $i$-band magnitude.
             The top panel shows the fraction of stars
             as a function of magnitude.
             the galaxy and star purity for TPC and BMC techniques.
             The bottom two panels show
             the galaxy and star purity for morphological separation
             and BMC technique.
             We use the same threshold values from 
             Figure~\ref{fig:purity_mag} in the bottom four panels}.
    \label{fig:purity_mag_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_z.eps}
    \caption{The same figures as Figure~\ref{fig:purity_mag},
             but as a function of photometric redshift estimate
             (photo-$z$).}
    \label{fig:purity_z}
  \end{minipage}
\end{figure*}

@@

\section{Conclusions}
  \label{section:conclusions}

\textcolor{blue}{begin new text}

We have presented and analyzed a novel star-galaxy classification framework
for combining classifications from the CFHTLenS project.
In particular, we use four independent star-galaxy classification techniques:
a morphological separation method;
\texttt{TPC}, a supervised machine learning technique
based on prediction trees and a random forest;
\texttt{SOMc}, an unsupervised machine leraning approach
based on self-organizing maps and a random atlas;
and \texttt{HB}, a hierarichical bayesian template-fitting method
that we have modified and parallized.
Both \texttt{TPC} and \texttt{SOMc} are currently availble within
a software package named
\texttt{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}.
Our implementaion of \texttt{HB} and \texttt{BMC},
as well as IPython notebooks that were used to
produce the results in this paper,
are available at \url{https://github.com/EdwardJKim/astroclass}.

\textcolor{blue}{end new text}

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[letterpaper,margin=0.65in]{geometry}
\usepackage[ansinew]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
\usepackage{amssymb}
\usepackage[normalem]{ulem}
\usepackage{natbib}
\usepackage{url}
\usepackage{float}

@@

\label{firstpage}
\begin{abstract}
\textcolor{blue}{
Star-galaxy separation in currently ongoing and future
large photometric surveys, such as DES and LSST,
will detect a vast number of unresolved galaxies at faint magnitudes.
There exist a variety of star-galaxy classification techniques,
each with their own strenths and weaknesses.
In this paper, we present a novel hybrid, ensemble classification
framework that combines and fully exploits different techniques
to produce a more robust classification.
To demonstrate this hybrid technique,
we combine a purely morphological classifier,
a supervised machine learning technique based on random forest,
an unsupervised method based on self-organizing maps,
and a heirarchical Bayeisan template fitting algorithm.
Any of these four methods could
easily be replaced by a similar method
or be extended to include other existing techniques.
Testing with CFHTLenS data,
we demonstrate that our combination technique improves
the overall performance over any individual classification method.
}

\end{abstract}

@@

%\begin{figure*}
%  \begin{minipage}[t]{0.49\linewidth}
%    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
%    \caption{Purity of morphological separation,
%             TPC, and BMC techniques in 
%             as a function of the $i$-band magnitude.
%             The top panel shows the histogram of objects in the test data set.
%             The second panel shows the fraction of stars
%             as a function of magnitude. The error bars are Poisson.
%             The bottom two panels compare
%             the galaxy and star purity for morphological separation, TPC,
%             and BMC techniques as a function of magnitude.
%             The error bars for completeness and purity are not shown for
%             clarity, but they are approximately 10 percent.}
%    \label{fig:purity_mag}
%  \end{minipage}
%  \hfill
%  \begin{minipage}[t]{0.49\linewidth}
%    \includegraphics[width=\textwidth]{figures/purity_z.eps}
%    \caption{The same figures as Figure~\ref{fig:purity_mag},
%             but as a function of photometric redshift estimate
%             (photo-$z$).}
%    \label{fig:purity_z}
%  \end{minipage}
%\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/purity_mag.eps}
  \caption{Purity of morphological separation,
           TPC, and BMC techniques in 
           as a function of the $i$-band magnitude.
           The top panel shows the histogram of objects in the test data set.
           The second panel shows the fraction of stars
           as a function of magnitude. The error bars are Poisson.
           The bottom two panels compare
           the galaxy and star purity for morphological separation, TPC,
           and BMC techniques as a function of magnitude.
           The error bars for completeness and purity are not shown for
           clarity, but they are approximately 10 percent.}
  \label{fig:purity_mag}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/purity_z.eps}
  \caption{The same figures as Figure~\ref{fig:purity_mag},
           but as a function of photometric redshift estimate
           (photo-$z$).}
  \label{fig:purity_z}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/purity_g-r.eps}
  \caption{The same figure as Figure~\ref{purity_mag},
           but as a function of $g-r$ color.}
  \label{fig:purity_g-r}
\end{figure}

\textcolor{blue}{In Figure~\ref{fig:purity_mag}, 
we compare the star and galaxy purity values
for morphological separation, TPC, and BMC techniques
as a function of $i$-band magnitude.
Although morphological separation shows a slightly better performance
at bright magnitudes $i < 21$,
BMC outperforms both morphological separation and TPC
at faint magnitudes $i > 21$.
As shown in Figure.~\ref{fig:purity_mag},
the histogram of distribution of objects peaks at $i \sim 22$,
and BMC therefore outperforms the morphological separation technique
for the majority of objects.
It is also clear that BMC outperforms TPC over all magnitudes.
BMC can presumably accomplish this by combining information from
all base classifiers,
\eg giving more weight to the morphological separation method
at bright magnitudes.}

\textcolor{blue}{We also show the star and galaxy purity values
of Table~\ref{table:metrics_all} as a function of
photometric redshift in Figure~\ref{fig:purity_z}.
Photo-$z$ is estimated with
the \texttt{BPZ} algorithm~\citep{benitez2000bayesian}
and provided with
the CFHTLenS photometric redshift catalogue~\citep{Hildebrandt2012}.
Although the morphological separation method outperformed BMC
at bright magnitudes in Figure.~\ref{fig:purity_mag},
it is clear that BMC outperforms
both morphological separation and TPC over all redshift range.}

\textcolor{blue}{In Figure.~\ref{fig:purity_g-r},
we present how the star and galaxy purity values vary
as a function of $g-r$ color.
It is again clear that BMC outperforms
both morphological separation and TPC over all $g-r$ color range.  }

\subsection{Heterogeneous Training}
  \label{section:poor_training}

It is very costly in terms of telescope time to
obtain a large sample spectroscopic observations
down to the limiting magnitude of a photometric sample.
Thus, we investigate the impact of training set quality
by analyzing in detail a more realistic case
where the training data is available
only for a small number of objects at bright magnitudes.
To emulate this scenario,
we only use objects that have spectroscopic labels
from the VVDS 0226-04 field (which is located within the CFHTLS W1 field)
and impose a magnitude cut of $i < 22.0$ in the training data,
leaving us a training set with 1,365 objects.
We apply the same four star-galaxy classification techniques
and four different combination methods,
and test them on the same test data.
Since the demographics of sources in the test data set
does not show the same distribution as the training data set,
this also serves as a test of the efficacy of heterogenous training data.

\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           as applied to the CFHTLenS data in the VVDS W1 field
           with a magnitude cut of $i < 22.0$.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
  \input{tables/metrics_cut.tex}
  \label{table:metrics_cut}
\end{table*}

We present in Table~\ref{table:metrics_cut}
the same six metrics for each method,
and highlight the best method for each metric.
Overall, the results obtained for the reduced data set are remarkable.
With poor training data, our data-driven methods,
\texttt{TPC} and \texttt{SOMc},
suffer a significant decrease in performance,
although they still in general perform better than 
the template fitting algorithm
\texttt{HB}.
The performance of our morphological separation method
does not depend on the training data,
and shows the best result in the MSE metric.
It also outperforms other base classifiers
in galaxy purity at galaxy completeness of 0.9964
and in star purity at star completeness of 0.7145.

@@

%% ROC curve for vvds i < 20
%\begin{figure}
%  \centering
%  \includegraphics[width=\linewidth]{figures/roc_cut.eps}
%  \caption{A magnified view of the ROC curve of
%           four star-galaxy individual classification approaches
%           (TPC, SOMc, HB, and morphology)
%           and the best-performing combination technique (BMC).
%           trained with the reduced data set.
%           We show only one of the four combination methods, BMC,
%           which has the best overall performance.}
%  \label{fig:roc_cut}
%\end{figure}

%In Figure~\ref{fig:roc_cut}, we plot the ROC curve
%of all four individual classification algorithm
%and our BMC technique for the poor training data case.
%We note that a binary classifier,
%such as our cut-based morphological separation,
%whose output is either zero or one,
%is a single point in the ROC space.
%We vary the half-light radius to create a smooth ROC curve
%for morphological separation.
%It is clear that our BMC technique dominates
%each individual classification techniques in the ROC space.

%% weights map for vvds i < 20

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/som_colors_cut.eps}
    \caption{The same figure as Figure~\ref{fig:som_colors_cut}
        but for the BMC technique applied to the four individual methods
        trained with the CFHTLenS data in the VVDS W1 field
        with $i < 22$.}
    \label{fig:som_colors_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/weights_cut.eps}
    \caption{The same figure as FIgure~\ref{fig:weights_cut}
        but for the BMC technique applied to the four invidiual methods
        trained with the CFHTLenS data in the VVDS W1 field
        with $i < 22$.}
    \label{fig:weights_cut}
  \end{minipage}
\end{figure*}

@@

%\begin{figure*}
%  \begin{minipage}[t]{0.49\linewidth}
%    \includegraphics[width=\textwidth]{figures/purity_mag_cut.eps}
%    \caption{Purity of morphological separation,
%             TPC, and BMC techniques in 
%             Table~\ref{table:metrics_vvds}
%             as a function of the $i$-band magnitude.
%             The top panel shows the fraction of stars
%             as a function of magnitude.
%             the galaxy and star purity for TPC and BMC techniques.
%             The bottom two panels show
%             the galaxy and star purity for morphological separation
%             and BMC technique.
%             We use the same threshold values from 
%             Figure~\ref{fig:purity_mag} in the bottom four panels}.
%    \label{fig:purity_mag_cut}
%  \end{minipage}
%  \hfill
%  \begin{minipage}[t]{0.49\linewidth}
%    \includegraphics[width=\textwidth]{figures/purity_z.eps}
%    \caption{The same figures as Figure~\ref{fig:purity_mag},
%             but as a function of photometric redshift estimate
%             (photo-$z$).}
%    \label{fig:purity_z}
%  \end{minipage}
%\end{figure*}

@@

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
    \caption{The classification performance metrics for TPC, HB, morphology,
      and BMC as applied to the CFHTLenS data in the VVDS field
      with various magnitude cuts.
      The top panel shows the number of sources in the training set
      at corresponding magnitude cuts.
      We show only one of the four combination methods, BMC,
      which has the best overall performance.}
    \label{fig:perform_mag_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
      $z_{\text{spec}}$ cuts.}
    \label{fig:perform_z_cut}
  \end{minipage}
\end{figure*}
\section{Conclusions}
  \label{section:conclusions}

\textcolor{blue}{We
have presented and analyzed a novel star-galaxy classification framework
for combining classifications from the CFHTLenS project.
In particular, we use four independent star-galaxy classification techniques:
a morphological separation method;
\texttt{TPC}, a supervised machine learning technique
based on prediction trees and a random forest;
\texttt{SOMc}, an unsupervised machine leraning approach
based on self-organizing maps and a random atlas;
and \texttt{HB}, a hierarichical bayesian template-fitting method
that we have modified and parallized.
Both \texttt{TPC} and \texttt{SOMc} are currently availble within
a software package named
\texttt{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}.
Our implementaion of \texttt{HB} and \texttt{BMC},
as well as IPython notebooks that were used to
produce the results in this paper,
are available at \url{https://github.com/EdwardJKim/astroclass}.  }
